---
title: "Variation in myoglobin content of skeletal muscle of seal species."
author: "Emma Rand"
output:
  pdf_document: default
  word_document: default
  html_document: default
bibliography: reference.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE)
```

```{r packages}
library(tidyverse)
library(Rmisc)
```

# Introduction
Aquatic and marine mammals are able to dive underwater for extended periods as a result of having a higher muscle  myoglobin concentration than terrestial mammals.
Seal species are known to vary in dive length. We investigated whether the concentration of myoglobin differed between  species.

# Methods
```{r import}
# Data import
# the data are organised in to three columns, one for each species. 
seal <- read.table("raw_data/seal.txt", header = TRUE)
```

```{r tidy}
# put the data in tidy format
seal2 <- gather(data = seal, species, myoglobin)

# write tidy data to processed
write.table(seal2, 
            "processed_data/seal2.txt",
            row.names = FALSE)
```


```{r summarise}
sealsummary <- summarySE(data = seal2,
                         measurevar = "myoglobin",
                         groupvars = "species")
```

```{r extract}
# extract values for inline reporting

highestmean <- max(sealsummary$myoglobin)
highestse <- sealsummary$se[sealsummary$myoglobin == highestmean]
highestspp <- sealsummary$species[sealsummary$myoglobin == highestmean]
lowestmean <- min(sealsummary$myoglobin)
lowestse <- sealsummary$se[sealsummary$myoglobin == lowestmean]
lowestspp <- sealsummary$species[sealsummary$myoglobin == lowestmean]

```

We measured the myoglobin content of the skeletal muscle of `r sealsummary$N[1]` individuals in each of three species.
We used R [@R-core] with tidyverse packages [@tidyverse] for all analyses.



```{r testing}
mod <- aov(data = seal2, myoglobin ~ species)

# extract values for inline reporting
res <- summary(mod)[[1]]
fval <- round(res$`F value`[1], 2)
df1 <- res$Df[1]
df2 <- res$Df[2]
p <- round(res$`Pr(>F)`[1], 3)
```

```{r posthoc, include=FALSE}
TukeyHSD(mod)
```

# Results
There is a significant difference in myoglobin concentration between species (_F_ = `r fval`; _d.f._=`r df1`, `r df2`; _p_ = `r p`). Post-hoc testing revealed that difference to be between the `r highestspp` seal with the highest myoglobin concentrations ($\bar{x} \pm s.e.$: `r round(highestmean,2)` $\pm$ `r round(highestse,2)` g Kg^-1^) and the `r lowestspp` seal with the lowest (`r round(lowestmean,2)` $\pm$ `r round(lowestse,2)` g Kg^-1^). See Figure 1.


```{r myo-fig, fig.height=4, fig.width=4, fig.cap="Figure 1 Mean Myoglobin content of skeletal muscle"}

ggplot(seal2, aes(x = species, y = myoglobin) ) +
  geom_jitter(width = .2, colour = "#8c8c8c") +
  geom_errorbar(data = sealsummary, 
                aes(ymin = myoglobin, 
                    ymax = myoglobin), 
                width =.2, size = 1 ) + 
  geom_errorbar(data = sealsummary, 
                aes(ymin = myoglobin - se, 
                    ymax = myoglobin + se), 
                width = .3 ) +
  ylab(expression("Myoglobin concentration g "*Kg^{-1})) +
  ylim(0, 75) +
  xlab("Seal species") +
  scale_x_discrete(labels = c("Bladdernose", 
                              "Harbour", 
                              "Weddell")) +
  theme(panel.background = element_rect(fill = "white")) +
  theme(axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"))


```


# References

